\documentclass[fleqn]{hw}

\usepackage{notes}
\usepackage{url}
\usepackage{graphicx}
\usepackage{amsmath}
\title{P3: Reinforcement Learning}
\duedate{}
\class{CS6300: Artificial Intelligence, Spring 2018}
\institute{University of Utah}
\author{Jake Pitkin}
% IF YOU'RE USING THIS .TEX FILE AS A TEMPLATE, PLEASE REPLACE
% The author WITH YOUR NAME AND UID.
% Replace the due date with anyone you worked with i.e. "Worked with: John McCarthy, Watson, & Hal-9000"
\begin{document}
\maketitle

\section*{Self Analysis}

\begin{enumerate}
	\item \textbf{What was the hardest part of the assignment for you?}
	
I found this project to be easy. I am comfortable enough with value iteration and q-learning from doing the written homework that implementing them was straightforward.
	\item \textbf{What was the easiest part of the assignment for you?}
	
I didn't hit any roadblocks when completing this project and was able to finish in a reasonable time.
	\item \textbf{What problem(s) helped further your understanding of the course material?}

Implementing q-learning helped me understand it a little better. The implementation is straightforward enough though that I would like to spend some more time with the theory behind q-learning.
	\item \textbf{Did you feel any problems were tedious and not helpful to your understanding of the
material?}

I didn't.
	\item \textbf{What other feedback do you have about this homework?}

Playing with the discounts and noise rates was a fun aspect that helped me understand them a little better.
\end{enumerate}


\end{document}
